---
title: "Report RetroMetInspector 🔎"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(readr)
library(DT)
library(tidyr)

final_rds = readRDS(snakemake@input[["final"]])
final2_rds = readRDS(snakemake@input[["final2"]])
ref_plots_dir = snakemake@input[["ref_plots"]]
custom_plots_dir = snakemake@input[["custom_plots"]]
met_window = snakemake@params[["met_window"]]
min_met_diff = snakemake@params[["min_met_diff"]]
min_reads = snakemake@params[["min_reads"]]
sample = snakemake@wildcards[["sample"]]
custom_subdir <- file.path(custom_plots_dir, "methylartist_plots")
cuantos=nrow(final2_rds)
```
# Transposable Elements (TE) with haplotype and met info for the sample: `r sample`

## General stadistics



### Average Genotype distribution for a total of `r cuantos` TEs
```{r echo=FALSE, fig.show='hold', fig.align='center'}
Final_met<- final2_rds
mis_colores <- terrain.colors(length(table(Final_met$Genotype)))
barplot(table(Final_met$Genotype), 
       main = "Genotype Distribution",
       ylab = "nº",
       xlab = "Genotype",
       col = mis_colores,
       legend.text = TRUE, 
       args.legend = list(x = "topright", bty = "n")) # "bty='n'" quita el recuadro de la leyenda
```


### Average meth-rate diff across all TE
```{r, include=FALSE}
#Final_met<-Final_met%>%filter(MetDiff_L>=0.05|MetDiff_R>=0.)
filtrado<-Final_met%>%filter(Genotype=="Het",
Insert_HP==1|Insert_HP==2)%>%mutate(
  hipHPmet=case_when(
    LeftHp1>LeftHp2|RightHp1>RightHp2~1,
    LeftHp2>LeftHp1|RightHp2>RightHp1~2
  ))
filtrado$Insert_HP=as.numeric(filtrado$Insert_HP)

Final<-filtrado%>%mutate(
  Matches=case_when(
    Insert_HP==hipHPmet~TRUE,
    Insert_HP!=hipHPmet~FALSE
))
MATCHES=Final%>%filter(Matches==TRUE)
MATCHES2 <- MATCHES %>%
  mutate(
    FLANKS = case_when(
      !is.na(MetDiff_L) & !is.na(MetDiff_R) ~ "BOTH",
      !is.na(MetDiff_L)                      ~ "LEFT",
      !is.na(MetDiff_R)                      ~ "RIGHT"
    )
  )
NOMATCHES=Final%>%filter(Matches==FALSE)


sum(MATCHES2$FLANKS=="LEFT")
A=MATCHES2[MATCHES2$FLANKS == "LEFT", ]
sum(MATCHES2$FLANKS=="LEFT")
```


```{r, echo=FALSE, fig.width=10, fig.height=5}
mis_colores <- terrain.colors(length(table(Final_met$MetDiff_L)))
par(mfrow = c(1, 2))
barplot(table(Final_met$MetDiff_L), 
        main = "Left-Flank meth diff",
        ylab = "nº",
        xlab = "Meth-Rate DIFF",
        col = mis_colores)

 mis_colores <- terrain.colors(length(table(Final_met$MetDiff_R)))
barplot(table(Final_met$MetDiff_R), 
        main = "Right-Flank meth diff",
        ylab = "nº",
        xlab = "Meth-Rate DIFF",
        col = mis_colores)
par(mfrow = c(1, 1))
```

### Analysis INFO:

**From the total of TEs (`r cuantos`) TEs 🧬:**

- **`r nrow(MATCHES2)`  TEs have been detected with "het" genotype and hypermethylation effect 🔴 **


- **A total of (`r nrow(NOMATCHES)`) TEs have been detected with "het" genotype and hypomethylation effect  🔵**


- **The remaining TEs  have been considered as "hom" genotype or contained NULL or NA values**


### TE families whose insertions in heterozygosity hypermethylate the allele

```{r echo=FALSE, fig.show='hold', fig.align='center'}
mis_colores <- terrain.colors(length(table(MATCHES2$Class)))

barplot(table(MATCHES2$Class), 
        main = "Family Distribution",
        ylab = "nº",
        xlab = "ETs-Families",
        col = mis_colores,
        legend.text = TRUE, 
        args.legend = list(x = "topleft", bty = "n")) # "bty='n'" quita el recuadro de la leyenda


```

### Average SVLEN of the TE whose insertion hypermethylates the allele

```{r echo=FALSE, fig.show='hold', fig.align='center'}
mis_colores <- terrain.colors(length(table(MATCHES2$SvLen)))

barplot(table(MATCHES2$SvLen), 
        main = "TE SvLen distribution",
        ylab = "nº",
        xlab = "SVLEN (bp)",
        col = mis_colores,
        legend.text = FALSE, 
        args.legend = list(x = "topleft", bty = "n")) # "bty='n'" quita el recuadro de la leyenda

```

### Meth-rate diff of the TE whose insertion hypermethylates the allele

```{r, echo=FALSE, fig.width=10, fig.height=5}
# Dividimos el espacio en 1 fila y 2 columnas
par(mfrow = c(1, 2))

# Gráfica 1
mis_colores_L <- terrain.colors(length(table(MATCHES2$MetDiff_L)))
barplot(table(MATCHES2$MetDiff_L), 
        main = "Left-Flank meth diff",
        col = mis_colores_L)

# Gráfica 2
mis_colores_R <- terrain.colors(length(table(MATCHES2$MetDiff_R)))
barplot(table(MATCHES2$MetDiff_R), 
        main = "Right-Flank meth diff",
        col = mis_colores_R)

# Es buena práctica resetear a 1 fila, 1 columna al terminar
par(mfrow = c(1, 1))


```

## Methylation analysis for a `r met_window`bp  (left-right) window


<details>
  <summary>Haz clic aquí para ver la tabla de todas las inserciones</summary>


```{r, echo=FALSE}
datatable(final2_rds) %>%
  # 1. ESTILO GLOBAL: Se aplica a TODA la tabla (columna 0)
  formatStyle(
    0, 
    target = "row", 
    textAlign = 'center', 
    fontFamily = 'Arial', 
    fontSize = '12px',
    backgroundColor = "#f5f5f5"
  ) %>%
  # 2. BLOQUES DE COLORES (tus reglas específicas)
  formatStyle(6:7, backgroundColor = "#e8f3ec",fontWeight = 'bold') %>%
  formatStyle(8:12, backgroundColor = "#e7f1fb") %>%
  formatStyle(13:17, backgroundColor = "#dbe9f6")%>%
  formatStyle(
    8, # La primera columna del bloque morado
    borderLeft = '2px solid white' 
  )%>%
  formatStyle(
    13, # La primera columna del bloque morado
    borderLeft = '2px solid white' 
  )%>%
formatStyle(
 6, # La primera columna del bloque morado
  borderLeft = '2px solid white' 
)%>% formatStyle(1,fontWeight = 'bold')
```
## Methylation analysis for a `r met_window`bp  (left-right) window and almost `r min_met_diff * 100`% of differences between them.  
<details>
  <summary>Haz clic aquí para ver la tabla de inserciones con filtro `r min_met_diff * 100`% </summary>
```{r, echo=FALSE}

datatable(final_rds) %>%
  # 1. ESTILO GLOBAL: Se aplica a TODA la tabla (columna 0)
  formatStyle(
    0, 
    target = "row", 
    textAlign = 'center', 
    fontFamily = 'Arial', 
    fontSize = '12px',
    backgroundColor = "#f5f5f5"
  ) %>%
  # 2. BLOQUES DE COLORES (tus reglas específicas)
  formatStyle(6:7, backgroundColor = "#e8f3ec",fontWeight = 'bold') %>%
  formatStyle(8:12, backgroundColor = "#e7f1fb") %>%
  formatStyle(13:17, backgroundColor = "#dbe9f6")%>%
  formatStyle(
    8, # La primera columna del bloque morado
    borderLeft = '2px solid white' 
  )%>%
  formatStyle(
    13, # La primera columna del bloque morado
    borderLeft = '2px solid white' 
  )%>%
formatStyle(
 6, # La primera columna del bloque morado
  borderLeft = '2px solid white' 
)%>% formatStyle(1,fontWeight = 'bold')
## Methylartist locus plots 
```


# TE Plots according the run parameters: `r min_reads` support reads and `r min_met_diff * 100`% diff methylation rate per each Haplotype.

## ***Flank plots***

This plots doesn't contain the meth-info of the insert.


⚠️ Notes: 

- Plot quality is related to the number of supporting reads and coverage.

```{r, results='asis', echo=FALSE}
ruta <- normalizePath(ref_plots_dir, winslash = "/", mustWork = TRUE)

lista <- list.files(
  ruta, full.names = TRUE,
  pattern = "\\.(png|jpg|jpeg)$",
  ignore.case = TRUE
)

for (i in lista) {
  i_abs <- normalizePath(i, winslash = "/", mustWork = FALSE)

  if (!file.exists(i_abs)) {
    cat(sprintf("<p style='color:red;'>No existe: %s</p>\n", i_abs))
    next
  }

  uri <- knitr::image_uri(i_abs)

  cat("<details>\n")
  cat(sprintf("<summary>📊 %s</summary>\n\n", basename(i_abs)))
  cat(sprintf("![](%s)\n\n", uri))
  cat("</details>\n\n")
}
```

### ***Custom plots***

This plots contain the meth-info of the insert.


⚠️ Notes: 

- Plot quality is related to the number of supporting reads and coverage.

```{r, results='asis', echo=FALSE}
ruta <- normalizePath(custom_subdir, winslash = "/", mustWork = TRUE)

lista <- list.files(
  ruta, full.names = TRUE,
  pattern = "\\.(png|jpg|jpeg)$",
  ignore.case = TRUE
)

for (i in lista) {
  i_abs <- normalizePath(i, winslash = "/", mustWork = FALSE)

  if (!file.exists(i_abs)) {
    cat(sprintf("<p style='color:red;'>No existe: %s</p>\n", i_abs))
    next
  }

  uri <- knitr::image_uri(i_abs)

  cat("<details>\n")
  cat(sprintf("<summary>📊 %s</summary>\n\n", basename(i_abs)))
  cat(sprintf("![](%s)\n\n", uri))
  cat("</details>\n\n")
}
```
