---
title: "RetroMetInspector Comparative Report 🔎 vs 🔎"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(readr)
library(DT)
library(tidyr)

samples<-snakemake@params[["samples"]]
met_window = snakemake@params[["met_window"]]
min_met_diff = snakemake@params[["min_met_diff"]]
min_reads = snakemake@params[["min_reads"]]
soi=snakemake@params$soi

#Cargamos los RDS finales
S1_ALL = readRDS(snakemake@input[["S1_ALL"]])
S2_ALL = readRDS(snakemake@input[["S2_ALL"]])
S1_filtered=readRDS(snakemake@input[["S1_filtered"]])
S2_filtered=readRDS(snakemake@input[["S2_filtered"]])
sample1=soi[1]
sample2=soi[2]
```



# Unshared Transposable Elements (TEs) detected between samples:`r sample1` and:`r sample2`


⚠️ Notes: 

- "Differential insertions between samples should be approached with caution. Variant callers such as **Sniffles2** or **cuteSV** can occasionally fail to detect a variant in a sample where it is actually present, thereby generating **false positives** in differential analysis."
- We recommend to manually check each TE using the igv_screenshot script aviable in igv folder.




- ** Columns with a blue background indicate  TE presence  🔵 **
- ** Columns with a red background indicate TE absence 🔴 **

## TE insertions without filtering parameters

### All TE insertions present in sample `r sample1`(S1) but not in sample `r sample2`(S2) with a `r met_window`bp  (left-right) meth window analysis.




<details>
  <summary>Haz clic aquí para ver la tabla de todas las inserciones</summary>

```{r, echo=FALSE}
S1_ALL$sample=NULL
datatable(S1_ALL) %>%
  formatStyle(
    0,
    target = "row",
    textAlign = 'center',
    fontFamily = 'Arial',
    fontSize = '12px',
    backgroundColor = "#f5f5f5"
  ) %>%
  formatStyle(6:7, backgroundColor = "#e8f3ec", fontWeight = 'bold') %>%
  formatStyle(8:12, backgroundColor = "#e7f1fb") %>%
  formatStyle(13:17, backgroundColor = "#dbe9f6") %>%
  formatStyle(18:23, backgroundColor = "#F08080") %>%
  formatStyle(23:27, backgroundColor = "#FF6A6A") %>%
  formatStyle(8,  borderLeft = '2px solid white') %>%
  formatStyle(13, borderLeft = '2px solid white') %>%
  formatStyle(6,  borderLeft = '2px solid white') %>%
  formatStyle(1, fontWeight = 'bold')
```



### All TE insertions present in sample `r sample2`(S2) but not in sample `r sample1`(S1) with a `r met_window`bp  (left-right) meth window analysis.


<details>
  <summary>Haz clic aquí para ver la tabla de todas las inserciones</summary>

```{r, echo=FALSE}
S2_ALL$sample=NULL
datatable(S2_ALL) %>%
  formatStyle(
    0,
    target = "row",
    textAlign = 'center',
    fontFamily = 'Arial',
    fontSize = '12px',
    backgroundColor = "#f5f5f5"
  ) %>%
  formatStyle(6:7, backgroundColor = "#e8f3ec", fontWeight = 'bold') %>%
  formatStyle(8:12, backgroundColor = "#e7f1fb") %>%
  formatStyle(13:17, backgroundColor = "#dbe9f6") %>%
  formatStyle(18:22, backgroundColor = "#F08080") %>%
  formatStyle(22:27, backgroundColor = "#FF6A6A") %>%
  formatStyle(8,  borderLeft = '2px solid white') %>%
  formatStyle(13, borderLeft = '2px solid white') %>%
  formatStyle(6,  borderLeft = '2px solid white') %>%
  formatStyle(1, fontWeight = 'bold')
```


## TE insertions with filtering parameters

### TE insertions present in sample `r sample1`(S1) but not in sample `r sample2`(S2) according the run parameters: r met_window`bp  (left-right) meth window analysis and `r min_met_diff * 100`% diff methylation rate per each Haplotype.



<details>
  <summary>Haz clic aquí para ver la tabla de todas las inserciones</summary>

```{r, echo=FALSE}
S1_filtered$sample=NULL
datatable(S1_filtered) %>%
  formatStyle(
    0,
    target = "row",
    textAlign = 'center',
    fontFamily = 'Arial',
    fontSize = '12px',
    backgroundColor = "#f5f5f5"
  ) %>%
  formatStyle(6:7, backgroundColor = "#e8f3ec", fontWeight = 'bold') %>%
  formatStyle(8:12, backgroundColor = "#e7f1fb") %>%
  formatStyle(13:17, backgroundColor = "#dbe9f6") %>%
  formatStyle(18:22, backgroundColor = "#F08080") %>%
  formatStyle(22:27, backgroundColor = "#FF6A6A") %>%
  formatStyle(8,  borderLeft = '2px solid white') %>%
  formatStyle(13, borderLeft = '2px solid white') %>%
  formatStyle(6,  borderLeft = '2px solid white') %>%
  formatStyle(1, fontWeight = 'bold')
```

### TE insertions present in sample `r sample2`(S2) but not in sample `r sample1`(S1) according the run parameters: r met_window`bp  (left-right) meth window analysis and `r min_met_diff * 100`% diff methylation rate per each Haplotype.



<details>
  <summary>Haz clic aquí para ver la tabla de todas las inserciones</summary>

```{r, echo=FALSE}
S2_filtered$sample=NULL
datatable(S2_filtered) %>%
  formatStyle(
    0,
    target = "row",
    textAlign = 'center',
    fontFamily = 'Arial',
    fontSize = '12px',
    backgroundColor = "#f5f5f5"
  ) %>%
  formatStyle(6:7, backgroundColor = "#e8f3ec", fontWeight = 'bold') %>%
  formatStyle(8:12, backgroundColor = "#e7f1fb") %>%
  formatStyle(13:17, backgroundColor = "#dbe9f6") %>%
  formatStyle(18:22, backgroundColor = "#F08080") %>%
  formatStyle(22:27, backgroundColor = "#FF6A6A") %>%
  formatStyle(8,  borderLeft = '2px solid white') %>%
  formatStyle(13, borderLeft = '2px solid white') %>%
  formatStyle(6,  borderLeft = '2px solid white') %>%
  formatStyle(1, fontWeight = 'bold')
```